Distribution-oblivious Programming with Dish

### 2019-11-18 pldi/draft.tex
Add spectrum figure
Add overview figure
Update intro

Distributed systems offer significant benefits over their centralized counterparts---for example, they can speed up expensive computations or process data that would not fit into any single machine.
% offer notable benefits over their centralized counterparts:
%   using partitioning and replication, they can process and store data with increased throughput and fault-tolerance.
% Yet, only a minority of developers, employed by the select few companies that deal with massive datasets, have the luxury of engineering software systems with distribution baked in from the start.
% The remaining majority starts by developing and deploying software in a centralized manner---that is, \emph{until} there is a significant change of requirements, such as a load increase.
Despite these benefits, their development remains different from and significantly more difficult than their centralized counterparts.
Whereas anyone can quickly stitch together a Bash script to compute on a single computer, 
  % domain-experts routinely glue scripts together to process and share data. % without the help of a computing expert.
   scaling out to multiple ones requires expert labor around ``point'' solutions with expensive setups, restricted programming interfaces, and exorbitant composition costs~\cite{taurus:14, dios:13, andromeda:15, pywren:17, futuredata:18, nefele:18}.

To understand this sharp contrast, consider calculating term frequencies over a set of input files:

@@

\noindent
Program $p_1$ creates a character stream, breaks it into words, transliterates to lower-case, sorts to group duplicates, summarizes them with a counter, sorts them in descending order, picks the top five results and writes them to a file \ttt{out}.
Combining several features, \unix makes small tasks easy to express;
  for a user with \emph{one} computer and small input, this pipeline takes a few seconds to compose and execute~\cite{bentley1986literate}.
% composition allows the composition of general primitives

@@

@@

For a user with many computers and larger inputs, however, developing $p_1$'s distributed equivalent requires a significant effort.
% The scope of such rewrites, and therefore the cost of manual effort, can vary considerably.
For simple pipelines that capture in restricted models of computation, this effort amounts to expressing the computation using the primitives provided by a big-data framework~\cite{mapreduce:08, ciel:11, spark:12, naiad:13} or domain-specific language~\cite{alvaro2011consistency, distal:13, meiklejohn2015lasp}---an unjustifiable cost for one-off pipelines that take a few minutes to compose (but are applied to large datasets).
More complex pipelines, such as the ones presented in later sections, require full-fledged distributed programming language~\cite{erlang:96, lopes1997d, acute:05, mace:07, cloudhaskell:11, ScalaLoci:18}. %  or a distributed operating system---\eg Plan9's \ttt{rc} shell.
In these cases, manual rewriting is expensive and can introduce new bugs, cascading changes, or divergence from legacy functionality.
Could the generation and execution of $p_1$'s distributed version be fully and correctly automated?

@@

The key insight behind \sys is that the language of the Unix shell already encodes stream processing, providing most of the information required for distributing a computation.
\sys builds on this insight with a careful study of the scalability properties of shell primitives and its commands, paired with a novel rewriting scheme that identifies maximal pipeline sub-expressions that are candidates for scale-out.
We term the result \emph{distribution-oblivious programming}---a series of techniques that fully automates program distribution while provably maintaining correct (non-distributed) semantics.

Key benefits include the ability to scale legacy computations onto
\sys converts legacy shell pipelines into their distributed equivalents fully automatically, offering $100\times$ improvements in performance without a single line of additional code or annotation.
Most importantly, \sys provides an architectural lesson:
  the current design shows that large-scale efforts in the distributed- and operating-system literature to provide a \unix-like distributed equivalent could have been simplified by a thin (but sophisticated) rewriting shim like the one \sys provides.

@@

@@

>>>>

Distributed systems offer significant benefits over their centralized counterparts---for example, they can speed up expensive computations or can process data that would not fit into any single machine.
% offer notable benefits over their centralized counterparts:
%   using partitioning and replication, they can process and store data with increased throughput and fault-tolerance.
% Yet, only a minority of developers, employed by the select few companies that deal with massive datasets, have the luxury of engineering software systems with distribution baked in from the start.
% The remaining majority starts by developing and deploying software in a centralized manner---that is, \emph{until} there is a significant change of requirements, such as a load increase.
Despite these benefits, their development remains different from and significantly more difficult than their centralized counterparts.
Whereas anyone can quickly stitch together a Bash script to compute on a single computer, 
  % domain-experts routinely glue scripts together to process and share data. % without the help of a computing expert.
   scaling out to multiple ones requires expert labor around ``point'' solutions with expensive setups, restricted programming interfaces, and exorbitant composition costs~\cite{taurus:14, dios:13, andromeda:15, pywren:17, futuredata:18, nefele:18}.

To understand this sharp contrast, consider a Bash pipeline calculating term frequencies over a set of inputs:

@@

\noindent
Program $p_1$ creates a character stream, breaks it into words, transliterates to lower-case, sorts to group duplicates, reduces duplicates to a counter, sorts counters in descending order, picks the top five results, and writes them to a file \ttt{out}.
Combining several features, \unix makes small tasks easy to express;
  for a user with \emph{one} computer and small input size, this pipeline takes a few seconds to compose and execute~\cite{bentley1986literate}.
% composition allows the composition of general primitives

@@

\begin{figure}[t]
\centering
\includegraphics[width=0.49\textwidth]{./figs/dish_spectrum.pdf}
\caption{
  \textbf{Cost of Manual Effort.}
	\sys sits at the automation end of the spectrum, automatically distributing shell pipelines while maintaining their correctness. A more complete picture is presented in the related-work section~\sx{rt}.
}
\label{fig:spectrum}
\end{figure}

@@

Unfortunately, developing $p_1$'s distributed equivalent requires a significant effort.
% For a user with many computers and larger inputs, 
% The scope of such rewrites, and therefore the cost of manual effort, can vary considerably.
For simple pipelines, fitting into restricted models of computation, this effort amounts to expressing the computation using the primitives provided by a big-data framework~\cite{mapreduce:08, ciel:11, spark:12, naiad:13} or domain-specific language~\cite{alvaro2011consistency, distal:13, meiklejohn2015lasp}---an unjustifiable cost for one-off pipelines that take a few minutes to compose (but are applied to large datasets).
More complex pipelines, such as the ones presented in later sections, would involve a full-fledged distributed programming language~\cite{erlang:96, lopes1997d, acute:05, mace:07, cloudhaskell:11, ScalaLoci:18}. %  or a distributed operating system---\eg Plan9's \ttt{rc} shell.
In both cases, manual rewriting is expensive and can introduce new bugs, cascading changes, and divergence from legacy functionality.
Could the generation and execution of $p_1$'s distributed version be fully and correctly automated?

@@

To answer this question, this paper presents \sys, a shell \todo{preprocessor} that transforms shell pipelines into their distributed equivalents enabling \emph{distribution-oblivious programming}.
The key insight behind \sys is that the language of the Unix shell already encodes stream processing, providing most of the information required to distribute a computation.
\sys builds on this insight with a careful study of the distributability properties of shell primitives and commands, a pipeline rewriting pass that identifies maximal sub-expressions as candidates for scale-out, and a late-bound command-prefixing scheme for deferring scheduling decisions to the distributed planner (\ie when runtime information is available).
A \sys-enabled $p_1$ will run \ttt{tr} and \ttt{uniq} in completely parallel streams, use a mostly-parallel tree of \ttt{sort}s, run \ttt{head} in one of the streams, and (trivially) merge streams in \ttt{out}.

The combination of automated transformations for program distribution with the ability to maintain correct non-distributed semantics results in several benefits.
% \sys converts legacy shell pipelines into their distributed equivalents fully automatically, offering $100\times$ improvements in performance without a single line of additional code or annotation.
First, shell users composing pipelines (or simply running legacy pipelines on massive datasets) can see scalability benefits without any manual effort---no need for \ttt{qsub}, SLURM, calls to \ttt{parallel}, or any manual rewriting.
Second, developers of new shell commands can use a concise a domain-specific language to express their distributability properties rather than using ad-hoc, command-specific flags such as {\tt t}  \#threads, \ttt{-t}, \ttt{-p} \etc
Most importantly, \sys provides an architectural lesson for system designers---namely that large-scale research efforts in the distributed- and operating-system literature to provide a \unix-like distributed equivalent could have been simplified by a thin (but sophisticated) rewriting shim like the one \sys provides.

@@

\begin{figure*}[t]
\centering
\includegraphics[width=0.49\textwidth]{./figs/dish_overview.pdf}
\caption{
  \textbf{Applying \sys to $p_1$.}
}
\label{fig:example}
\end{figure*}

@@

\begin{figure}[t]
\centering
\includegraphics[width=0.45\textwidth]{./figs/dish_ex1.pdf}
\label{fig:example}
\end{figure}
